---
title: "Machine Learning Assignment"
author: "Julian Oehling"
date: "14 Februar 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The purpose of this document is to develop a machine learning model which predicts how well a fitness excersise was performed by participants based on the data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.

First, the data is read in:
```{r reading data}
library(caret)
train <- read.csv(file = "E:/Coursera/MachineLearning/pml-training.csv", header = TRUE, na.strings=c(""," ","NA"))
quiz <- read.csv(file = "E:/Coursera/MachineLearning/pml-testing.csv", header = TRUE, na.strings=c(""," ","NA"))
```

## Data cleaning

A first inspection shows, that many columns contain mostly NAs. They are removed in the next step.

```{r remving NAs}
train <- train[,colSums(!is.na(train)) > (nrow(train)*0.9)]
```

This reduces the number of columns from 160 to 60. A few of them contain information which is not helpful, for example a unique id number. They are excluded as well.

```{r removing useless information}
train <- train[,-1]
train <- train[,-(2:4)]
```

## Data splitting

To estimate the out of sample error when the model has been built, 10% of the data are removed from the training data set and put into a test dataset.

```{r splitting}
library(caret)
trainIndex <- createDataPartition(train$classe, p = .9, 
                                  list = FALSE, 
                                  times = 1)
test <- train[-trainIndex,]
train <- train[trainIndex,]
```


## Model building

Now, a model can be built in order to predict the class. In order to create a class predictor a random forest approach is chosen. 10-fold cross validation is used and principal component analysis for the 55 variables.

```{r ML, warning=FALSE, echo=FALSE, eval=FALSE}
ctrl <- trainControl(method = "repeatedcv", repeats = 10)
fit <- train(classe ~., data= train, method="parRF", trControl = ctrl, preProcess = "pca")
```
Due to the very long calculation time (several hours!) this step was not reperfomed when finally knitting the document (see: eval=FALSE) but the result loaded below. The code is otherwise unaltered.


## Model analysis

Finally, let's have a look at the model.

```{r model}
load("E:/Coursera/MachineLearning/fit.Rdata")
fit$results
```


In the last step, the out of sample error is estaimated by using the model on the previously created test data.

```{r oose}
confusionMatrix(predict(fit, newdata = test), test$classe)
```

It can be seen,that the expected out of sample error is very small.

## Predictin on the quiz data

As required 20 predictions are made for the week 4 quiz.

```{r quiz}
predict(fit, newdata = quiz)
```

